The Evolution of Heavy-Ion Physics: A Data-Driven Analysis of Quark Matter Conferences

### 2025-03-24 QM/paper.tex
all are fine except for figures/QM_keyword_analysis.pdf

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/QM_keyword_analysis.pdf}
\caption{Analysis of research topics in Quark Matter conferences from 2011 to 2023. The left panel shows the evolution of key physics concepts as percentages of total presentations, while the right panel displays trends in experimental facilities mentions. This visualization reveals quantitative trends in research focus over time, tracking the rise and fall of specific physics topics and the changing emphasis on different experimental facilities.}
\label{fig:keywords}
\end{figure}

Figure~\ref{fig:keywords} presents an analysis of research topics across Quark Matter conferences from 2011 to 2023. This visualization provides quantitative trend analysis to offer insights into the field's evolution:

@@

@@

@@

The temporal trends in the right panel show some notable patterns. While European representation has remained relatively stable, there are visible fluctuations that often correlate with conference locations. Conferences held in Europe typically show elevated European participation, and similar regional effects are observable for North American and Asian venues.

Asian participation shows a gradual upward trend across the analyzed period, reflecting the growing strength of heavy-ion physics research in countries like China, Japan, and India. This trend suggests a slow but steady diversification of the field's geographical base.

>>>>

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/QM_keyword_analysis.pdf}
\caption{Analysis of research topics in Quark Matter conferences from 2011 to 2025. The left panel shows the evolution of key physics concepts as percentages of total presentations, while the right panel displays trends in experimental facilities mentions. This visualization reveals quantitative trends in research focus over time, tracking the rise and fall of specific physics topics and the changing emphasis on different experimental facilities.}
\label{fig:keywords}
\end{figure}

Figure~\ref{fig:keywords} presents an analysis of research topics across Quark Matter conferences from 2011 to 2025. This visualization provides quantitative trend analysis to offer insights into the field's evolution:

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/keyword_trends.pdf}
\caption{Detailed evolution of key research topics in Quark Matter conferences from 2011 to 2025. This visualization tracks the frequency of important physics concepts in presentation titles over time, organized into related thematic clusters. Each line represents the percentage of presentations mentioning a specific research area, allowing for direct comparison of how different topics have risen or fallen in prominence. The trends highlight both enduring themes like QGP properties and emerging areas such as small systems and ultra-peripheral collisions (UPC), providing a more nuanced view of topic evolution compared to the aggregate analysis.}
\label{fig:keyword_trends}
\end{figure}

Figure~\ref{fig:keyword_trends} provides a more detailed examination of research topic evolution, extending the analysis in Figure~\ref{fig:keywords}. This visualization offers several key insights:

The clustered organization of topics reveals how related research areas have co-evolved over time. For example, we can observe the correlation between the rise of machine learning techniques and the increased focus on Bayesian analysis methods, reflecting the field's growing adoption of advanced statistical approaches for data interpretation.

The temporal trends also highlight the impact of major experimental milestones on research focus. Peaks in certain topics often follow significant data-taking periods at RHIC or the LHC, demonstrating how the field responds to new experimental capabilities and results.

The visualization captures emerging research frontiers that have gained momentum in recent years. Topics such as small system collective effects, which investigate whether QGP-like behavior occurs in proton-nucleus collisions, show notable growth trends that suggest developing research directions.

By tracking the relative prominence of different topics over time, this analysis provides a quantitative basis for understanding how the intellectual landscape of heavy-ion physics has evolved. This information can help researchers identify both established areas with sustained interest and emerging opportunities for novel contributions.

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/regional_diversity_by_year.pdf}
\caption{Detailed view of regional representation at Quark Matter conferences from 2011 to 2025. This stacked area chart shows the percentage of total presentations from each geographical region for each conference year. The visualization reveals both persistent patterns and year-to-year fluctuations in international participation. North America and Europe consistently contribute the largest proportion of presentations, while Asia shows gradual growth in representation. The minimal presence of research from South America, Africa, and Oceania highlights opportunities for increased geographical inclusivity in future conferences. Annual variations often correlate with conference host locations, with regions typically showing elevated participation when hosting.}
\label{fig:regional_diversity_by_year}
\end{figure}

Figure~\ref{fig:regional_diversity_by_year} expands upon the regional analysis with a more detailed year-by-year breakdown. Several important patterns are evident:

The temporal trends show notable fluctuations that often correlate with conference locations. Conferences held in Europe typically show elevated European participation, and similar regional effects are observable for North American and Asian venues. This "host region effect" appears to be a consistent factor in participation patterns.

Asian participation shows a gradual upward trend across the analyzed period, reflecting the growing strength of heavy-ion physics research in countries like China, Japan, and India. This trend suggests a slow but steady diversification of the field's geographical base, though predominance from traditional research centers remains strong.

The visualization clearly identifies years with anomalous participation patterns, such as during the COVID-19 pandemic, when travel restrictions significantly impacted international representation at conferences. These disruptions highlight the sensitivity of international scientific participation to global events.

@@
